Push-Pull Signal-Function Functional Reactive Programming

### 2012-11-29 Docs/IFL2012/design.tex
1st pass of section 2.4

@@

Signal functions are constructed from combinators, which are primitive signal
functions and operations to combine these primitives. These combinators are
grouped as basic signal functions, lifting operations for pure functions,
routing, reactivity, feedback, event processing, joining, and time dependence.

@@

The routing combinators are used to combine signal functions, and to re-arrange
and discard components of signal vectors in order to connect signal functions.

Only those routing combinators which modify or combine signal functions
({\tt (>>>)}, {\tt first}, {\tt second}) are reactive (may replace themselves in
response to events, and then only if they inherit their reactivity from the
signal function(s) which are inputs to combinator. The rest do not
react to or modify the input in any way, except to re-arrange it, copy it, or
discard it altogether.

@@

This combinator provides decoupling for signals
(the input signal is the output signal at the previous time-step)
but not events (event occurrences are supplied to the combinator immediately).
This means that the programmer has the responsibility to ensure that feedback
does not generate an infinite sequence of events in a single time-step.

The {\tt filter} and {\tt filterList} combinators are stateless signal functions
which apply a function to each event occurrence. The {\tt filter} combinator
expects an option-typed output, while the {\tt list} combinator expects a list
output. If the function yields a non-empty output, one or more output occurrences
are produced with the elements of that output. If the output is empty, there
are no output occurrences.

The {\tt accumulate} and {\tt accumulateList} functions are similar, except that
they maintain a state value which is replaced after each input event occurrence
by an additional output from the function, and supplied to the function in
addition to the value of each event occurrence.

The joining combinators provide the ability to combine two event
streams, two signals, or a signal and an event stream.

@@

The {\tt combineSignals} combinator applies a binary function pointwise to two
signals, and produces the result of this application as a third signal. The
combining function is necessary because we will have two input samples at each
time step, and must produce exactly one output sample.

The {\tt capture} combinator adds the last-sampled value of a signal at the time
of an event occurrence to that event occurrence.

These three combinators together provide the ability to combine elements of
signal vectors. By combining these combinators, along with the {\tt cancelLeft}
and {\tt cancelRight} routing combinators, arbitrary signal vectors can be
reduced.

A set of combinators are provided for making use of time-dependence in a signal
function. These combinators allow the modification of signals and events with
respect to time, and the observation of the current time value.

The simplest time-dependent combinator is {\tt time}, which simply outputs
the time since it began to be evaluated. This does not necessarily correspond to
the time since the global signal function began to be evaluated, since the
signal function in which the {\tt time} combinator is used may have been
introduced through {\tt switch}.

@@

>>>>

% (set-fill-column 100)

@@

%% Signal functions are constructed from combinators, which are primitive signal
%% functions and operations to combine these primitives. 
\new{TimeFlies includes basic signal functions and a library of signal function combinators, falling
  into these categories:}
% These combinators are grouped as basic signal functions, 
lifting operations from pure functions to signal functions,
routing, reactivity, feedback, event processing, joining, and time dependence.
%
\rn{Eek, I get thrown off by a definition of combinators which is so broad as to include everything.  I
  like the definition here 
   \url{http://stackoverflow.com/tags/combinators/info} : 
   {\em A combinator is a higher-order function that uses only function application and earlier defined
    combinators to define a result from its arguments.}  Does everything in the above list meet that
   definition?}

@@

The routing combinators are used to combine signal functions, and to re-arrange
and discard components of signal vectors in order to connect signal functions.
%
Only those routing combinators which modify or combine signal functions
({\tt (>>>)}, {\tt first}, {\tt second}) are reactive (\ie{} may replace themselves in
response to events, and then only if they inherit their reactivity from the
signal function(s) which are inputs to combinator. The rest do not
react to or modify the input in any way, except to re-arrange it, copy it, or
discard it altogether.

@@

This combinator provides decoupling for signals
(the input signal is the output signal at the previous time-step)
but not events (event occurrences are supplied to the combinator immediately).
This means that the programmer has the responsibility to ensure that feedback
does not generate an infinite sequence of events in a single time-step.
\rn{Presumably that is different from previous formulations of FRP?  Because ... there was no
  notion of multiple events in a sampling interval?}

The {\tt filter} and {\tt filterList} combinators are stateless signal functions
which apply a function to each event occurrence. The {\tt filter} combinator
expects an option-typed output, while the {\tt list} combinator expects a list
output. \rn{and input right?} If the function yields a non-empty output, one or more output occurrences
are produced with the elements of that output. If the output is empty, there
are no output occurrences.

The {\tt accumulate} and {\tt accumulateList} functions are similar, except that
they maintain a state value which is replaced after each input event occurrence
by an additional output from the function, and supplied to the function in
addition to the value of each event occurrence. \rn{Murky prose definition...}

The joining combinators provide the ability to combine two event
streams, two signals, or a signal and an event stream. 
\rn{How does that work temporally?  For event/event is this a zip
  that matches up the left events w/ the right events?}

@@

The {\tt combineSignals} combinator applies a binary function pointwise to two
signals, and produces the result of this application as a third signal. The
combining function is necessary because we will have two input samples at each
time step, and must produce exactly one output sample.
\rn{Wait that sounded like it was covered by join?}

The {\tt capture} combinator adds the last-sampled value of a signal at the time
of an event occurrence to that event occurrence.
\rn{Again, what is a signal-event {\bf join} if not that?}

These three combinators \rn{Which?} together provide the ability to combine elements of
signal vectors. By combining these combinators, along with the {\tt cancelLeft}
and {\tt cancelRight} routing combinators, arbitrary signal vectors can be
reduced.

A set of combinators are provided for making use of time-dependence in a signal
function. These combinators allow the modification of signals and events with
respect to time, and the observation of the current time value.
%
The simplest time-dependent combinator is {\tt time}, which simply outputs
the time since it began to be evaluated. This does not necessarily correspond to
the time since the global signal function began to be evaluated, since the
signal function in which the {\tt time} combinator is used may have been
introduced through {\tt switch}.

\rn{If that's the simplest what are the more complex ones?  If not worth mentioning it seems easier to just
    describe {\bf time}.}

@@

\todo{What's novel here?  After this section is de-listified need to emphasize that.}
